---
title: "Introduction to atctools Package"
author: "Ville Langén"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Overview of the atctools Package Functions and Datasets}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(atctools)
```


# Introduction

Welcome to the `atctools` package vignette! This vignette provides an overview of the package functionality, demonstrates the usage of key functions, and explains the example datasets included.

## Overview

The `atctools` package provides tools for ATC (Anatomical Therapeutic Chemical) code matching and data integrity checks for drug names. It includes functions for matching drug names to ATC codes with and without stripping trailing characters. 

The atctools package comprises:

- Two main functions for drug name matching and ATC code assignment:
   - **`drug_match`**: Matches drug names to ATC codes without stripping trailing characters.
   - **`drug_match_strip`**: Matches drug names to ATC codes, stripping trailing characters such as "mg", "g", etc.
   - **`atc2drug`**: Matches ATC codes to drug names.
- Three example datasets to illustrate package usage:
   - **`example_drug_data`**: Drug names without milligrams.
   - **`example_drug_data_with_mg`**: Drug names including milligrams.
   - **`example_reference_data`**: Reference data with ATC codes and corresponding drug names.
   - **`example_atc_data`**: ATC codes for testing the atc2drug function, including valid, non-valid, and NA values.



## Example Datasets

### atctools::example_drug_data


This dataset contains example drug names without milligrams.

```{r}
example_drug_data
```

### atctools::example_drug_data_with_mg

This dataset contains example drug names including milligrams.

```{r}
example_drug_data_with_mg
```

### atctools::example_reference_data

This dataset is a reference dataset containing example drug names and corresponding ATC codes for drug matching. Please ensure that:

- The first column (in this example: `drug_name`) contains drug names.
- The second column (in this example: `ATC_code`) contains ATC codes.

```{r}
example_reference_data
```

### atctools::example_atc_data

This dataset provides example ATC codes for testing the `atc2drug` function. It includes valid ATC codes, non-valid codes, and `NA` values to demonstrate how the function handles different types of input.

```{r}
example_atc_data
```

### atctools::example_atc_data_for_validation

This dataset contains a mix of valid, invalid, and unknown ATC codes across two columns (`atc_1`, `atc_2`) for testing the validation functions.

```{r}
example_atc_data_for_validation
```


## Functions

### atctools::validate_atc()

The `validate_atc()` function checks whether each ATC code in a specified column is format-valid based on the official ATC code structure. It adds a new column (e.g., `atc_1_invalid`) next to the specified column indicating whether the code is invalid (`1`) or valid (`0`).

ATC codes must be either 5 or 7 characters long and follow the expected letter-number-letter-letter-number-number structure (for 7-character codes).

```{r}
# Validate ATC codes by format
validate_output <- validate_atc(example_atc_data_for_validation, "atc_1")

# View result
validate_output
```

### atctools::validate_atc_with_reference()

The `validate_atc_with_reference()` function checks if each ATC code in a specified column exists in a reference list of valid codes (e.g., from `example_reference_data`). This is useful for ensuring codes are not just structurally valid but also known and accepted.

```{r}
# Validate ATC codes against a reference list
validate_by_ref_output <- validate_atc_with_reference(example_atc_data_for_validation, "atc_2", example_reference_data)

# View result
validate_by_ref_output
```

#### Example of Combined Output

Here’s how you might run both validation functions in sequence and examine the combined results:

```{r}
output_combined <- example_atc_data_for_validation %>%
  validate_atc("atc_1") %>%
  validate_atc_with_reference("atc_2", example_reference_data)

output_combined
```

The flags:

- `atc_1_invalid`: Format-based invalidity
- `atc_2_invalid`: Not found in reference list


### atctools::drug_match()

The `drug_match` function matches drug names in a dataset to corresponding ATC codes using a reference dataset. It does not strip trailing characters from drug names.

Usage:
```{r}
library(atctools)

# Perform drug matching
output_data1 <- drug_match(example_drug_data, "drug_name", example_reference_data)

# Display output data
output_data1
```

### atctools::drug_match_strip()

The `drug_match_strip` function matches drug names in a dataset to corresponding ATC codes, stripping trailing characters (e.g., "mg", "g") from drug names before matching.

Usage:

```{r}
# Perform drug matching with trailing characters stripped
output_data2 <- drug_match_strip(example_drug_data_with_mg, "drug_name", example_reference_data)

# Display output data
output_data2
```


### atctools::atc2drug()

The `atc2drug` function matches ATC codes to drug names based on a reference dataset. It handles valid ATC codes, ignores invalid ATC codes, and returns `NA` for non-matching codes.

Usage:

```{r}
# Perform ATC code to drug name matching
output_data3 <- atc2drug(example_atc_data, "ATC_code", example_reference_data)

# Display output data
output_data3
```



## Flag Meaning in the Output Data of `example_drug_data` and `example_drug_data_with_mg`

The `drug_name_flag` column in the output dataset indicates whether there is a prefix mismatch between the drug name and its matched ATC code.

- 1: Indicates a mismatch where the first three characters of the drug name do not match the first three characters of the ATC code.
- 0: Indicates a match where the first three characters of the drug name match the first three characters of the ATC code.


## Conclusion

This vignette provides an overview of how to use the atctools package for drug name matching and understanding the flag column in the output. For more details on each function and additional functionality, refer to the package documentation and examples.

