---
title: "Introduction to atctools Package"
author: "Ville Langén"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{introduction_to_my_package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(atctools)
```


# Introduction

Welcome to the `atctools` package vignette! This vignette provides an overview of the package functionality, demonstrates the usage of key functions, and explains the example datasets included.

## Overview

The `atctools` package provides tools for ATC (Anatomical Therapeutic Chemical) code matching and data integrity checks for drug names. It includes functions for matching drug names to ATC codes with and without stripping trailing characters. 

The atctools package comprises:

- Two main functions for drug name matching and ATC code assignment:
   - **`drug_match`**: Matches drug names to ATC codes without stripping trailing characters.
   - **`drug_match_strip`**: Matches drug names to ATC codes, stripping trailing characters such as "mg", "g", etc.
- Three example datasets to illustrate package usage:
   - **`example_drug_data`**: Drug names without milligrams.
   - **`example_drug_data_with_mg`**: Drug names including milligrams.
   - **`example_reference_data`**: Reference data with ATC codes and corresponding drug names.



## Example Datasets

### `example_drug_data`


This dataset contains example drug names without milligrams.

```r
example_drug_data
```
### `example_drug_data_with_mg`

This dataset contains example drug names including milligrams.

```r
example_drug_data_with_mg
```

### `example_reference_data``

This dataset is a reference dataset containing example drug names and corresponding ATC codes for drug matching. Please ensure that:

- The first column (in this example: `drug_name`) contains drug names.
- The second column (in this example: `ATC_code`) contains ATC codes.

```r
example_reference_data
```




## Functions

### `drug_match`

The `drug_match` function matches drug names in a dataset to corresponding ATC codes using a reference dataset. It does not strip trailing characters from drug names.

Usage:
```r
library(atctools)

# Perform drug matching
output_data1 <- drug_match(example_drug_data, "drug_name", example_reference_data)

# Display output data
output_data1
```

### `drug_match_strip`


The `drug_match_strip` function matches drug names in a dataset to corresponding ATC codes, stripping trailing characters (e.g., "mg", "g") from drug names before matching.

Usage:

```r
# Perform drug matching with trailing characters stripped
output_data2 <- drug_match_strip(example_drug_data_with_mg, "drug_name", example_reference_data)

# Display output data
output_data2
```

## Flag Meaning

The `drug_name_flag` column in the output dataset indicates whether there is a prefix mismatch between the drug name and its matched ATC code.

- 1: Indicates a mismatch where the first three characters of the drug name do not match the first three characters of the ATC code.
- 0: Indicates a match where the first three characters of the drug name match the first three characters of the ATC code.

## Conclusion

This vignette provides an overview of how to use the atctools package for drug name matching and understanding the flag column in the output. For more details on each function and additional functionality, refer to the package documentation and examples.

